Understanding both enhanced and impaired learning with enhanced plasticity: a saturation hypothesis (supplementary material)

### 2016-06-10 MHCpaper/MHCsupp.tex
removed unnecessary stuff from supp

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
\documentclass[10pt]{article}
%\usepackage[printfigures]{figcaps}
%\usepackage[nolists,nomarkers]{endfloat}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
\hypersetup{colorlinks,linkcolor=blue,citecolor=teal,urlcolor=teal}
\input{sl_graphics_preamble.tex}
\graphicspath{{"Figures/"}}
%
\usepackage{array}
\newcounter{tablerow}[table]
\newcommand{\tablerowautorefname}{row}
\newenvironment{tabularn}[1]{\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
%\usepackage{etoolbox}
%\AtBeginEnvironment{tabular}{\setcounter{tablerow}{0}}
\renewcommand{\figureautorefname}{Supplementary Figure}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\I}{\mathbf{I}}
%vec of ones
%equilibrium distribution
\newcommand{\pr}{\mathbf{p}}
\newcommand{\eq}{\pr^\infty}
%other symbols
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\norm}{_0}
\newcommand{\inc}{_{\text{inc}}}
\newcommand{\dec}{_{\text{dec}}}
\newcommand{\incdec}{_{\text{inc/dec}}}
\newcommand{\wt}{_{\text{WT}}}
\newcommand{\ko}{_{\text{K$^\mathrm{b}$D$^\mathrm{b}$-/-}}}
\newcommand{\KO}{K$^\mathrm{b}$D$^{\mathrm{b}-/-}$}
%\newcommand{\ko}{_{\text{MHC}}}
%\newcommand{\KO}{MHC}
\newcommand{\tpre}{t_{\text{pre}}}
\newcommand{\ttrain}{t_{\text{train}}}
%\newcommand{\lmax}{_{\text{max}}}
%\newcommand{\lmin}{_{\text{min}}}
\newcommand{\modelfig}[1][a]{Figure 4#1 of the main text}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{Understanding both enhanced and impaired learning with enhanced plasticity: a saturation hypothesis (supplementary material)}
%
%% Author List:
%%
%\author{An Author$^a$, Another Author$^b$ and Yet Another Author$^c$\\
%%
%\small{\emph{$^a$ Address 1}}\\
%\small{\emph{$^b$ Address 2}}\\
%\small{\emph{$^c$ Address 3}}\\
%}
\date{}
\DeclareCaptionLabelFormat{supp}{\textbf{#1 S#2}}
\captionsetup{labelformat=supp}

@@

We make the following assumptions:
\begin{itemize}
  \item There are $N$ identical synapses with $M$ internal functional states.
  \item The states of different synapses are independent of each other.
  \item The synapses that are eligible for plasticity are chosen randomly.
  \item Candidate potentiating/depressing plasticity event timings are distributed as Poisson processes with rates $r\potdep$.
  \item Potentiation and depression are described by Markov processes with transition probabilities $\M\potdep$.
  \item The synaptic weights of the internal states are given by the column vector $\w$. This can only take values in a finite range that we can shift to $\pm1$. Most of the models will only use the two extreme values.
\end{itemize}

In practice, the overall rate of candidate plasticity events will only set the timescale, and only the relative rates will matter.
With this in mind, we define
%
\begin{equation}\label{eq:fpotdep}
  r = r\pot + r\dep,
  \qquad
  f\pot = \frac{r\pot}{r},
  \qquad
  f\dep = \frac{r\dep}{r}.
\end{equation}
%
The quantities $f\potdep$ measure the fraction of candidate plasticity events that are potentiating/depressing.

@@

For models with only two possible synaptic weights, the distribution of synaptic weights is completely described by the mean, $\pr(t)\w$.

The \KO\ mice have a lower threshold for depression.
We can model this by changing $\M\dep\wt$ to $\M\dep\ko$, which should have larger off-diagonal matrix elements.

@@

For the serial, multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation and depression in the wild-type as well as potentiation in the \KO\ models.
We will use a larger value for $q$ for depression in the \KO\ models.

For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation and depression in the wild-type as well as potentiation in the \KO\ models.
We will use a larger value for $x$ for depression in the \KO\ models.

@@

\begin{figure}
 \begin{center}
 \parbox{0.8\linewidth}{%
 \begin{myenuma}
  \item\aligntop{\includegraphics[height=1.7cm]{serial.svg}}\label{fig:serial_model}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}\hspace{0.5cm}\\[1cm]
  \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate_model}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=2.5cm]{cascade.svg}}\label{fig:cascade_model}
 \end{myenuma}
 }
 \end{center}
  \caption[Transition probabilities for different models]{Transition probabilities for different models.
  Potentiation induces transitions indicated by orange arrows, depression indicated by blue arrows.
  States of strong/weak synaptic weight indicated by orange/blue circles.
  (\ref{fig:serial_model}) In the serial model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
  The synaptic weight takes only two values, $\pm1$.
  (\ref{fig:binary_model}) The two-state model is parameterised by the two transition probabilities.
  (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised  by these two values.
  The synaptic weight varies linearly in the interval $[-1,1]$.
  (\ref{fig:cascade_model}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}) and synaptic weight takes only two values.
  } \label{fig:models}
\end{figure}

@@

Suppose that there is some resource required for potentiation/depression that is shared between $P$ synapses and is depleted as more synapses are potentiated/depressed and replenished when this is reversed.
We can avoid going beyond the independent synapse model by modelling this pool of synapse as a single compound synapse.

@@

The Markov process is lumpable \wrt this partition of states (see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time).
The transition probabilities between lumps $i$ and $j$ is computed by choosing one state from lump $i$ and summing the transition probabilities to all states in lump $j$.
The result must be the same for all states in lump $i$.

@@

This model is parameterised by the range of values, $q\in[q\lmin,q\lmax]$, for potentiation and depression.
We will use the same values for potentiation and depression in the wild-type as well as potentiation in the \KO\ models.
We will use larger values for depression in the \KO\ models.
In the main text, we consider a version of this model that has resource depletion for depression only, so that potentiation transition probabilities are unaffected by the number of potentiated synapses.
This is done by setting $q\pot\lmax=q\pot\lmin$.
The values of these parameters are listed in \autoref{fig:sim_results}\ref{tab:params}.

\begin{figure}
 \begin{center}
  \aligntop{\includegraphics[height=4cm]{pooled.svg}}
 \end{center}
  \caption[The pooled resource model]{The pooled resource model.
  Several two-state synapses share resources that are required for potentiation and depression.
  These resources are depleted as more synapses are potentiated or depressed.
  This pool of synapses can be modelled as one compound synapse.} \label{fig:pooled_model}
\end{figure}

@@

This model is similar to the multistate model (see \autoref{fig:models}\ref{fig:multistate_model})), as it only has transitions between adjacent states and a linearly varying synaptic weight.
However, like the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade_model}), the transition probabilities decay exponentially away from the central transition.
More precisely:
%
\begin{equation}\label{eq:nonunidef}
    \M\pot_{ii+1} = \M\dep_{i+1i} =  x^{\abs{\frac{M+1}{2}-i}+\half},
      \quad i = 1 \ldots M-1,
\end{equation}
%
with all other off-diagonal elements equal to zero.
The diagonal elements are chosen so that the rows sum to one.

@@

\begin{figure}
 \begin{center}
  \aligntop{\includegraphics[height=3cm]{multistate_nonuni.svg}}
 \end{center}
  \caption[The non-uniform multistate model]{The non-uniform multistate model.
  Similar to the multistate model, (\autoref{fig:models}\ref{fig:multistate_model}), the synaptic weight varies linearly in the interval $[-1,1]$, but the transition probabilities between adjacent states decays exponentially away from the central transition for both potentiation and depression.} \label{fig:nonuni_model}
\end{figure}

@@

Training the animal will not change the internal dynamics of a synapse under potentiation or depression.
It will change the environment, which will lead to a change in how often potentiation and depression occur.
It could be manifested in a change in which synapses are potentiated/depressed, but this could not be captured in this type of model.
We will model this by changing $f\potdep$, leaving $\M\potdep$ unchanged.
This approach was used to model motor learning in \cite{Smith2006savings}.

@@

For the most part, we set $f\dep\norm=\half$, $f\dep\inc=f\dep\norm+\Delta f$ and $f\dep\dec=f\dep\norm-\Delta f$, with $\Delta f>0$.
We use the same values for wild-type and \KO\ models for the reasons discussed above.
We will mostly treat gain-increase and decrease symmetrically, but this is not necessary.
We could adjust $r$ to keep $r\pot$ unchanged, if so desired, but this would only change the overall timescale and would not affect any of the qualitative comparisons that we are concerned with here.
The values of these parameters are listed in \autoref{fig:sim_results}\ref{tab:params}.

@@

\begin{figure}[p]
\begin{myenuma}
  \begin{tabular}{l@{\hspace{0.05\linewidth}}l@{\hspace{0.05\linewidth}}l}
      \item\label{fig:seial_sim}
      \aligntop{\includegraphics[width=0.25\linewidth]{serial.svg}}
      &
      \item\label{fig:binary_sim}
      \hspace{0.05\linewidth}\aligntop{\includegraphics[width=0.125\linewidth]{binary.svg}}
      &
      \item\label{fig:multistate_sim}
      \aligntop{\includegraphics[width=0.25\linewidth]{multistate.svg}}
      \\[1.5cm]
      \aligntop{\includegraphics[width=0.3\linewidth]{serial_med_learnS.eps}}
      &
      \aligntop{\includegraphics[width=0.3\linewidth]{binary_learnS.eps}}
      &
      \aligntop{\includegraphics[width=0.3\linewidth]{multistate_lin_learnS.eps}}
      \\[5.5cm]
      \item\label{fig:pooled_sim}
      \aligntop{\includegraphics[width=0.3\linewidth]{pooled_deponly.svg}}
      &
      \item\label{fig:cascade_sim}
      \aligntop{\includegraphics[width=0.25\linewidth]{cascade.svg}}
      &
      \item\label{fig:nonuni_sim}
      \aligntop{\includegraphics[width=0.25\linewidth]{multistate_nonuni.svg}}
      \\[2cm]
      \aligntop{\includegraphics[width=0.3\linewidth]{pooled_deponly_learnS.eps}}
      &
      \aligntop{\includegraphics[width=0.3\linewidth]{cascade_long_learnS.eps}}
      &
      \aligntop{\includegraphics[width=0.3\linewidth]{nonuni_learnS.eps}}
  \end{tabular}

  \vspace{1cm}\item\label{tab:params}\aligntop{
  \begin{tabularn}{|l|c|c|c|c|c|c|}
    \cline{1-7}
    Model & \# states & pot & WT dep & \KO\ dep & $\Delta f$ & $r\tpre$ \\
    \cline{1-7}
    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.3  & 20  &\label{tr:multistate_med} \\
    Two-state     & 2  & $q=0.1$  & $q=0.1$  & $q=0.2$  & 0.1  & 5   &\label{tr:binary} \\
    Multistate    & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.3  & 5   &\label{tr:multistate_lin} \\
    Pooled res. & 7  & $q=0.008$        & $q\in[0.0006,0.6]$  & $q\in[0.001,1]$
                                          & 0.4 & 20 &\label{tr:pooled_deponly}\\
    Cascade       & 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 100 &\label{tr:cascade_long} \\
    Non-uni. MS& 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 150 &\label{tr:nonuni} \\
    \cline{1-7}
  \end{tabularn}}
  \end{myenuma}
  \caption[Simulation results for various models]{(\ref{fig:seial_sim}-\ref{fig:nonuni_sim}) Simulation results for various models, showing decrease in mean synaptic weight over time during gain increase training from the normal state (solid) and after gain-decrease pre-training (dashed) for wild-type (black) and \KO\ (red) models. (\ref{tab:params}) Parameters used for simulations.}\label{fig:sim_results}
\end{figure}

@@

First, consider the wild-type, for which $q\pot=q\dep=q$.
Without pre-training:
%
\begin{equation}\label{eq:multiWTnopre}
  \Phi = \frac{2{\Delta f}q}{M}.
\end{equation}
%
%where it's worth remembering that $\Delta f<0$.
With pre-training:
%
\begin{equation}\label{eq:multiWTpre}
\begin{aligned}
  \Phi &= 16(\Delta f)^2q \, \frac{(1+2\Delta f)^{M/2-1} (1-2\Delta f)^{M/2-1}}
          {(1+2\Delta f)^M - (1-2\Delta f)^M} \\
       &= \frac{4{\Delta f}\,q}{M} + \CO(\Delta f)^2.
\end{aligned}
\end{equation}
%
So, we see that pre-training will speed up learning when $\Delta f$ is small.
%, as seen in \autoref{fig:multistate_weak}\ref{fig:multistate_weak_learnS} (black curves).
On the other hand, if $\Delta f$ is close to $\half$, pre-training will initially slow down learning a lot due to the factor of $(1-2\Delta f)^{M/2-1}$.
%, as seen in \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}.

@@

Now, consider the \KO\ models, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
Without pre-training:
%
\begin{equation}\label{eq:multiKOnopre}
  \Phi = 2{\Delta f} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M}.
\end{equation}
%
This will be smaller than \eqref{eq:multiWTnopre} if $\beta<\beta^*(M)$, where $\beta^*(M)$ is defined as the value at which they are equal.
This function is plotted in \autoref{fig:multistate_star}\ref{fig:multistate_betastar}, where we can see that it approaches 1 rapidly as we increase $M$.

@@

In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f \in [\Delta f^*(1,M),\Delta f^*(\beta,M)]$ and look at intermediate time-scales, we will see that the \KO\ models learn slower than wild-type without pre-training, and that pre-training speeds up learning in the \KO\ models but slows it down in the wild-type.
%However, in this case pre-training proceeds much faster in the \KO\ models than wild-type (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learn} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}, lower curves), which is \emph{not} seen in the experiment.

@@

We find that
%
\begin{equation}\label{eq:binaryflux}
\begin{aligned}
  \Phi &= \frac{q\pot q\dep \Delta f}{f\pot q\pot + f\dep q\dep}, &\qquad
  \pdiff{\Phi}{q\dep} &= \frac{f\pot (q\pot)^2 \Delta f}{(f\pot q\pot + f\dep q\dep)^2}, \\ &&
  \pdiff{\Phi}{f\pot} &= \frac{(f\pot\inc q\pot + f\dep\inc q\dep) q\pot q\dep}{(f\pot q\pot + f\dep q\dep)^2}.
\end{aligned}
\end{equation}
%
The flux is a monotonically increasing function of $q\dep$ and $f\pot$.
Therefore, enhancing plasticity in this way or pre-training can only enhance the initial learning rate, unlike the experimental results shown in \modelfig.

>>>>

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
\documentclass[10pt]{article}
%\usepackage[printfigures]{figcaps}
%\usepackage[nolists,nomarkers]{endfloat}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
\hypersetup{colorlinks,linkcolor=blue,citecolor=teal,urlcolor=teal}
\input{sl_graphics_preamble.tex}
\graphicspath{{"Figures/"}}
%
\usepackage{array}
%\newcounter{tablerow}[table]
%\newcommand{\tablerowautorefname}{row}
%\newenvironment{tabularn}[1]{\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
%\usepackage{etoolbox}
%\AtBeginEnvironment{tabular}{\setcounter{tablerow}{0}}
\renewcommand{\figureautorefname}{Supplementary Figure}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\I}{\mathbf{I}}
%vec of ones
%equilibrium distribution
\newcommand{\pr}{\mathbf{p}}
\newcommand{\eq}{\pr^\infty}
%other symbols
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\norm}{_0}
\newcommand{\inc}{_{\text{inc}}}
\newcommand{\dec}{_{\text{dec}}}
\newcommand{\incdec}{_{\text{inc/dec}}}
\newcommand{\wt}{_{\text{WT}}}
\newcommand{\ko}{_{\text{K$^\mathrm{b}$D$^\mathrm{b}$-/-}}}
\newcommand{\KO}{K$^\mathrm{b}$D$^{\mathrm{b}-/-}$}
%\newcommand{\ko}{_{\text{MHC}}}
%\newcommand{\KO}{MHC}
\newcommand{\tpre}{t_{\text{pre}}}
\newcommand{\ttrain}{t_{\text{train}}}
%\newcommand{\lmax}{_{\text{max}}}
%\newcommand{\lmin}{_{\text{min}}}
\newcommand{\modelfig}[1][a]{Figure 4#1 of the main text}
\newcommand{\datafig}[1][b]{Figure 3#1 of the main text}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{Understanding both enhanced and impaired learning with enhanced plasticity: a saturation hypothesis (supplementary material)}
%
%% Author List:
%%
%\author{An Author$^a$, Another Author$^b$ and Yet Another Author$^c$\\
%%
%\small{\emph{$^a$ Address 1}}\\
%\small{\emph{$^b$ Address 2}}\\
%\small{\emph{$^c$ Address 3}}\\
%}
\date{}
\DeclareCaptionLabelFormat{supp}{\textbf{#1 S#2}}
\captionsetup{labelformat=supp}

@@

We make the following assumptions:
\begin{itemize}
  \item There synapses are identical and have $M$ internal functional states, where the states of different synapses are independent of each other.
%  \item The synapses that are eligible for plasticity are chosen randomly.
  \item Candidate potentiating/depressing plasticity event timings are distributed as Poisson processes with rates $r\potdep$.
  \item Candidate potentiation and depression events are described by Markov processes with transition probabilities $\M\potdep$.
  \item The synaptic weights of the internal states are given by the column vector $\w$. This can only take values in a finite range that we can shift to $\pm1$. Most of the models will only use the two extreme values.
\end{itemize}

The overall rate of candidate plasticity events will only set the timescale, and only the relative rates will matter.
With this in mind, we define
%
\begin{equation}\label{eq:fpotdep}
  r = r\pot + r\dep,
  \qquad
  f\pot = \frac{r\pot}{r},
  \qquad
  f\dep = \frac{r\dep}{r},
\end{equation}
%
where $f\potdep$ is the fraction of candidate plasticity events that are potentiating/depressing.

@@

%For models with only two possible synaptic weights, the distribution of synaptic weights is completely described by the mean, $\pr(t)\w$.

The \KO\ mice have a lower threshold for depression.
We can model this by changing $\M\dep\wt$ to $\M\dep\ko$, which has larger off-diagonal matrix elements.

@@

For the serial, multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation in the wild-type and \KO\ models.
We will use a larger value for $q$ for depression in the \KO\ models than in the wild-type.

For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation in the wild-type and \KO\ models.
We will use a larger value for $x$ for depression in the \KO\ models than in the wild-type.

@@

\begin{figure}
 \begin{center}
   \begin{myenuma}
  \begin{tabular}{l@{\hspace{0.05\linewidth}}l}
    \item\aligntop{\includegraphics[height=1.7cm]{serial.svg}}\label{fig:serial_model}&
    \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}\\[1.7cm]
    \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate_model}&
    \item\aligntop{\includegraphics[height=2.5cm]{cascade.svg}}\label{fig:cascade_model}\\[2.5cm]
    \item\aligntop{\includegraphics[height=2.3cm]{pooled.svg}}\label{fig:pooled_model}&
    \item\aligntop{\includegraphics[height=2.3cm]{multistate_nonuni.svg}}\label{fig:nonuni_model}
  \end{tabular}
   \end{myenuma}
 \end{center}
  \caption[Transition probabilities for different models]{Transition probabilities for different models.
  Potentiation induces transitions indicated by orange arrows, depression indicated by blue arrows.
  States of strong/weak synaptic weight indicated by orange/blue circles.
  (\ref{fig:serial_model}) In the serial model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
  The synaptic weight takes only two values, $\pm1$.
  (\ref{fig:binary_model}) The two-state model is parameterised by the two transition probabilities.
  (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised  by these two values.
  The synaptic weight varies linearly in the interval $[-1,1]$.
  (\ref{fig:cascade_model}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}) and synaptic weight takes only two values.
  (\ref{fig:pooled_model}) In the pooled resource model,
  Several two-state synapses share resources that are required for potentiation and depression.
  These resources are depleted as more synapses are potentiated or depressed.
  This pool of synapses can be modelled as one compound synapse.
  (\ref{fig:nonuni_model}) In the non-uniform multistate model
  the synaptic weight varies linearly in the interval $[-1,1]$, similar to the multistate model, but the transition probabilities between adjacent states decays exponentially away from the central transition for both potentiation and depression.
  } \label{fig:models}
\end{figure}

@@

Suppose that there is some resource required for potentiation/depression that is shared between $P$ synapses and is depleted as more synapses are potentiated/depressed and replenished when this is reversed, as shown in \autoref{fig:models}\ref{fig:pooled_model}.
We can avoid going beyond the independent synapse model by modelling this pool of synapse as a single compound synapse.

@@

The Markov process is lumpable \wrt this partition of states (see \cite{kemeny1960finite,burke1958markovian,Ball1993Lumpability}).
The transition probabilities between lumps $i$ and $j$ is computed by choosing one state from lump $i$ and summing the transition probabilities to all states in lump $j$,
which must be the same for all states in lump $i$.

@@

This model is parameterised by the range of values, $q\in[q\lmin,q\lmax]$, for potentiation and depression.
We will use the same values for potentiation in the wild-type asnd \KO\ models.
We will use larger values for depression in the \KO\ models than in the wild-type.
we consider a version of this model that has resource depletion for depression only, so that potentiation transition probabilities are unaffected by the number of potentiated synapses.
This is done by setting $q\pot\lmax=q\pot\lmin$.
The values of these parameters are listed in \autoref{fig:sim_results}\ref{tab:params}.

%\begin{figure}
% \begin{center}
%  \aligntop{\includegraphics[height=4cm]{pooled.svg}}
% \end{center}
%  \caption[The pooled resource model]{The pooled resource model.
%  Several two-state synapses share resources that are required for potentiation and depression.
%  These resources are depleted as more synapses are potentiated or depressed.
%  This pool of synapses can be modelled as one compound synapse.} \label{fig:pooled_model}
%\end{figure}

@@

This model, shown in \autoref{fig:models}\ref{fig:nonuni_model}, is similar to the multistate model (see \autoref{fig:models}\ref{fig:multistate_model})), as it only has transitions between adjacent states and a linearly varying synaptic weight.
However, like the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade_model}), the transition probabilities decay exponentially away from the central transition.
More precisely:
%
\begin{equation}\label{eq:nonunidef}
    \M\pot_{ii+1} = \M\dep_{i+1i} =  x^{\abs{\frac{M+1}{2}-i}+\half},
      \quad i = 1 \ldots M-1,
\end{equation}
%
with all other off-diagonal elements equal to zero.
The diagonal elements are chosen so that the rows sum to one.

@@

%\begin{figure}
% \begin{center}
%  \aligntop{\includegraphics[height=3cm]{multistate_nonuni.svg}}
% \end{center}
%  \caption[The non-uniform multistate model]{The non-uniform multistate model.
%  Similar to the multistate model, (\autoref{fig:models}\ref{fig:multistate_model}), the synaptic weight varies linearly in the interval $[-1,1]$, but the transition probabilities between adjacent states decays exponentially away from the central transition for both potentiation and depression.} \label{fig:nonuni_model}
%\end{figure}

@@

Training the animal will not change the internal dynamics of a synapse under potentiation or depression.
It will change the environment, which will lead to a change in how often potentiation and depression occur.
%It could be manifested in a change in which synapses are potentiated/depressed, but this could not be captured in this type of model.
We will model this by changing $f\potdep$, leaving $\M\potdep$ unchanged.
This approach was used to model motor learning in \cite{Smith2006savings}.

@@

For the most part, we set $f\dep\norm=\half$, $f\dep\inc=f\dep\norm+\Delta f$ and $f\dep\dec=f\dep\norm-\Delta f$, with $\Delta f>0$.
We use the same values for wild-type and \KO\ models for the reasons discussed above.
%We will mostly treat gain-increase and decrease symmetrically, but this is not necessary.
We could adjust $r$ to keep $r\pot$ unchanged, if so desired, but this would only change the overall timescale and would not affect any of the qualitative comparisons that we are concerned with here.
The values of these parameters are listed in \autoref{fig:sim_results}\ref{tab:params}.

@@

\begin{figure}[p]
\begin{myenuma}
  \begin{tabular}{lll}
      \item\label{fig:seial_sim}
      \aligntop{\includegraphics[width=0.24\linewidth]{serial.svg}}
      &
      \item\label{fig:binary_sim}
      \hspace{0.05\linewidth}\aligntop{\includegraphics[width=0.12\linewidth]{binary.svg}}
      &
      \item\label{fig:multistate_sim}
      \aligntop{\includegraphics[width=0.24\linewidth]{multistate.svg}}
      \\[1.3cm]
      \aligntop{\includegraphics[width=0.29\linewidth]{serial_fit_learnS.eps}}
      &
      \aligntop{\includegraphics[width=0.29\linewidth]{binary_learnS.eps}}
      &
      \aligntop{\includegraphics[width=0.29\linewidth]{multistate_lin_learnS.eps}}
      \\[3.5cm]
      \item\label{fig:pooled_sim}
      \aligntop{\includegraphics[width=0.29\linewidth]{pooled_deponly.svg}}
      &
      \item\label{fig:cascade_sim}
      \aligntop{\includegraphics[width=0.24\linewidth]{cascade.svg}}
      &
      \item\label{fig:nonuni_sim}
      \aligntop{\includegraphics[width=0.24\linewidth]{multistate_nonuni.svg}}
      \\[1.8cm]
      \aligntop{\includegraphics[width=0.29\linewidth]{pooled_deponly_learnS.eps}}
      &
      \aligntop{\includegraphics[width=0.29\linewidth]{cascade_fit_learnS.eps}}
      &
      \aligntop{\includegraphics[width=0.29\linewidth]{nonuni_fit_learnS.eps}}
  \end{tabular}

  \vspace{0.5cm}
  \item\label{tab:params}\aligntop{
  \begin{tabular}{|l|c|c|c|c|c|c|c|c|}
    \cline{1-9}
    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
    Model & \#  & \multicolumn{3}{c|}{Plasticity parameter} & \multicolumn{3}{c|}{$f\dep$} & $r\tpre$ \\
    \cline{3-8}
    & states & pot & WT dep & {\footnotesize\KO} dep & base & inc & dec & {} \\

    \cline{1-9}
    Serial   & 10 & $0.12$  & $0.14$  & $0.2$  & 0.5 & 0.89 & 0.11  & 100  \\%&\label{tr:serial_fit} \\
    Two-state     & 2  & $0.1$  & $0.1$  & $0.2$  & 0.5 & 0.6 & 0.4  & 5   \\%&\label{tr:binary} \\
    Multistate    & 10 & $0.3$  & $0.3$  & $0.4$  & 0.5 & 0.8 & 0.2  & 5   \\%&\label{tr:multistate_lin} \\
    Pooled res.\ & 7  & $0.008$        & $[0.0006,0.6]$  & $[0.001,1]$
                                          & 0.5 & 0.9 & 0.1 & 20 \\%&\label{tr:pooled_deponly}\\
    Cascade  & 14 & $0.386$  & $0.398$  & $0.466$  & 0.522 & 0.63 & 0.002  & 200  \\%&\label{tr:cascade_fit} \\
    Non-uni.\ & 12 & $0.4$    & $0.4$    & $0.53$   & 0.5 & 0.7 & 0.1 & 500  \\%&\label{tr:nonuni_fit} \\
    \cline{1-9}
  \end{tabular}}
  \end{myenuma}
  \caption[Simulation results for various models]{(\ref{fig:seial_sim}-\ref{fig:nonuni_sim}) Simulation results for various models, showing decrease in mean synaptic weight over time during gain increase training from the normal state (solid) and after gain-decrease pre-training (dashed) for wild-type (black) and \KO\ (red) models. (\ref{tab:params}) Parameters used for simulations. For the serial, two state and multistate models, the plasticity parameter listed is the transition probability between adjacent states. For the cascade and non-uniform multistate models, the plasticity parameter is the ratio of adjacent transition probabilities.}\label{fig:sim_results}
\end{figure}

@@

First, consider the wild-type.
Without pre-training:
%
\begin{equation}\label{eq:multiWTnopre}
  \Phi = \frac{2{\Delta f}q}{M}.
\end{equation}
%
%where it's worth remembering that $\Delta f<0$.
With pre-training:
%
\begin{equation}\label{eq:multiWTpre}
\begin{aligned}
  \Phi &= 16(\Delta f)^2q \, \frac{(1+2\Delta f)^{M/2-1} (1-2\Delta f)^{M/2-1}}
          {(1+2\Delta f)^M - (1-2\Delta f)^M} \\
       &= \frac{4{\Delta f}\,q}{M} + \CO(\Delta f)^2.
\end{aligned}
\end{equation}
%
So, we see that pre-training will speed up learning when $\Delta f$ is small.
%, as seen in \autoref{fig:multistate_weak}\ref{fig:multistate_weak_learnS} (black curves).
On the other hand, if $\Delta f$ is close to $\half$, pre-training will initially slow down learning a lot due to the factor of $(1-2\Delta f)^{M/2-1}$.
%, as seen in \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}.

@@

Now, consider the \KO\ models, for which we define $\beta=q\pot/q\dep<1$, and $q\pot=q$.
Without pre-training:
%
\begin{equation}\label{eq:multiKOnopre}
  \Phi = 2{\Delta f} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M}.
\end{equation}
%
This will be smaller than \eqref{eq:multiWTnopre} if $\beta<\beta^*(M)$, where $\beta^*(M)$ is defined as the value at which they are equal.
This function is plotted in \autoref{fig:multistate_star}\ref{fig:multistate_betastar}, where we can see that it approaches 1 rapidly as we increase $M$.

@@

In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f \in [\Delta f^*(1,M),\Delta f^*(\beta,M)]$, we will see that the \KO\ models learn slower than wild-type without pre-training, and that pre-training speeds up learning in the \KO\ models but slows it down in the wild-type.
%However, in this case pre-training proceeds much faster in the \KO\ models than wild-type (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learn} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}, lower curves), which is \emph{not} seen in the experiment.

@@

We find that
%
\begin{equation}\label{eq:binaryflux}
  \Phi = \frac{q\pot q\dep \Delta f}{\lambda}, \qquad
  \pdiff{\Phi}{q\dep} = \frac{f\pot (q\pot)^2 \Delta f}{\lambda^2}, \qquad
  \pdiff{\Phi}{f\pot} = \frac{\lambda\inc q\pot q\dep}{\lambda^2}.
\end{equation}
%
The flux is a monotonically increasing function of $q\dep$ and $f\pot$.
Therefore, enhancing plasticity in this way or pre-training can only enhance the initial learning rate, unlike the experimental results shown in \modelfig.

We have only looked at the initial learning rate here, but for the experiments we looked at the learning after a finite period of time.
As the total change will be larger for the wild-type than \KO\ models, it must overtake eventually.
If this happens early enough (\ie between the first and second data point in \datafig), the initial period where the \KO\ models learns faster than wild-type would not be seen in the experiment.
However, the timescale for this crossover would be similar to the timescale of the exponentials, which is at least as long as the timescale for saturation of learning.
In \datafig\ we see that this timescale is longer than the gaps between successive measurements, so we would not miss any initial period of enhanced learning in the \KO\ mice if this were the correct model.
In addition, the models without pre-training can never catch up with those with pre-training, and this is sufficient to rule out this model.
